Supplementary Material

### 2019-05-01 eDriversMS-SI-v2.rmd
latest modifications to manuscript, including frontiers formatting and supervisor comments

<!--
cd phdproj/edrivers/edriversms
rmarkdown::render('./eDriversMS-SI-v2.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS-SI.rmd', 'html_document')
-->

@@

Table S1. List of drivers currently available on *eDrivers* along with their respective accronym used in the figures in the supplementary material.

@@

We used $\Omega_{A}$ saturation state to characterize ocean acidification in the
bottom waters of the St. Lawrence. The data were collected [...]. $\Omega_{A}$,
pH, dissolved oxygen ($O_2$) and pressure were sampled at 117 stations in the
Estuary and Gulf of St. Lawrence in the summer of 2018. The majority of sampling
stations were located in the Southern Gulf ($n = 84$), compared to the Estuary
and Northern Gulf ($n = 33$).

@@

We explored the correlation between $\Omega_{A}$ and $O_2$ in the data and
found no significant correlation at the scale of the whole St. Lawrence (*p-value*:
0.08; $R^2$: 0.03). Nevertheless, a correlation
between $\Omega_A$ and $O_2$ exists in the Northern St. Lawrence (*p-value*:
< 0.01; $R^2$: 0.49), but not in the Southern St. Lawrence (*p-value*: 0.85;
$R^2$: < 0.01). Furthermore, the correlation observed in the north between $\Omega_A$
and $O_2$ is especially strong when deep Channels (*Deep* model: *p-value*:
< 0.01; $O_2$: 0.20; $R^2$: 0.97) are separated from shallower regions
(*Shallow* model: *p-value*: < 0.01; $O_2$: 0.07; $R^2$: 0.96), with $\Omega_A$
increasing more steeply with increasing $O_2$ in deep Channels. This stronger
correlation is likely caused by the lack of gas exchange with the atmosphere
or the surface mixed layer and the importance of metabolic $CO_2$ in deep
waters, some of which may have mean ages of up to 20 years [@mucci2011]. In the
Northern Gulf, the acidification is following
depth and distance from the Cabot Strait along the Laurentian, Anticosti and
Esquiman Channels due to the progressive oxygen depletion and metabolic carbon
dioxide accumulation [@mucci2011]. There are likely different dynamics at play
in the Southern Gulf, such as the accumulation of metabolic carbon dioxide
produced by rapid oxic degradation of organic matter in the sandy and permeable
sediments of the Magdalen Shelf [*personal communication?*]. <!-- see with Michel -->

@@

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentration in the water
column from the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2018] and interpolated oxygen % saturation using cokriging with depth as a
covariable, as done in [@dutil2011]. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

@@

The data used to characterize hypoxia uses oxygen saturation in % and come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] surveys in late spring and fall, as well as from annual
DFO’s multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. We provide a brief summary of data and methods to
characterize hypoxia in this document. For more details, refer to @blais2018.

Oxygen concentration is measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on the CTD. The probe is calibrated against seawater
samples that are analyzed by Winkler titration on every cast
[see @blais2018 for calibration procedure]. The data used is the last depth
sampled on the CTD profile, which typically corresponds to ~10m above the
bottom. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation taking into account the salinity and temperature at the selected
depth.

We interpolated oxygen saturation using cokriging with depth as a covariable,
as done in [@dutil2011]. Bathymetric data come from @dutil2011 and have a
100 $km^2$ resolution.   According to @diaz1995, severe hypoxia can be observed
when dissolved oxygen falls below 2 $ml$ $L^{-1}$, which corresponds to 62.5
$\mu mol$ $L^{-1}$ and ~20% saturation. This is considered as the level
necessary to maintain most animal life. Similarly, @chabot2008 studied the
effects of hypoxia on the energy budget of cod in the St. Lawrence and found
that behavioural effects began manifesting below 70% oxygen saturation and that
survival becomes jeopardised below ~20% saturation. We used these observations
to create an index of hypoxia using an inverted logistic curve that slowly
increases below 70% oxygen saturation, increases acutely as it approaches the
20% oxygen saturation threshold and reaches its peak intensity between 20 and
30% oxygen saturation:

@@

The surface layer is characterized using sea surface temperature (SST) monthly
composites from Advanced Very High Resolution Radiometer (AVHRR) satellite
images obtained from the National Oceanic and Atmospheric Administration (NOAA)
and European Organisation for the Exploitation of Meteorological Satellites
(EUMETSAT). Images used are from DFO’s Maurice Lamontagne Institute at a 1 km
resolution from 1985-2013 and from DFO’s Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5 km resolution since 2014. Monthly
anomalies were constructed as the difference between monthly averages and the
1985-2010 climatological mean for each month.

@@

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level. We therefore had
to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence.

@@

The data comes from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use glogally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

@@

As in @halpern2008b and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statisticscanada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modelled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic polllution.

@@

The information provided on this expert map on the 5 toxins [@bates2019].
was georeferenced and transformed as vecctorized objects.
We calculated a toxic algae risk ($T$) index for each cell ($x$) in the 1 $km^2$
study grid. For each toxin ($t$), a value of 1 was attributed to all grid cells
overlapping with areas identified at risk on the expert map and a value of 0.5
for grid cells overlapping with areas where ASP and DSP toxins were observed
without exceeding legal thresholds. The value for all 5 toxins was them summed
for all grid cells:

@@

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities in the
St. Lawrence. The data we used spans 6 years from 2010 to 2015 and details
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targetted species and a total of 53 caught species in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008b and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

@@

In order to characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multipled the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008b; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worlwide. Data used spanned 2003 to 2011.

@@

The data used to characterize invasive species risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008b and @halpern2015.

Cargo volume was used as a proxy of invasion risk under the assumption that
risk of invasion is proportional to tonnes of goods transferred through ports.
Cargo throughput in metric tonnes for the year 2011 was accessed through a
variety of sources [see supplementary material in @halpern2015 for more details]
and cross-matched with entries in the World Port Index database (WPI; available
from the National Geospatial-Intelligence Agency). A gap-filling procedure
using linear regression and sets of predictors related to port volume and
available in the WPI dataset was then applied to the WPI dataset to predict
missing cargo volume entries. Finally, volume data was distributed in marine
environments adjacent to ports using a diffusive plume model with an exponential
decay function that set the maximum spread distance to approximately 1000 km.
The plume model was then clipped to areas less than 60 m deep, as
invasives species are more likely to invade shallow areas.

@@

The data used to characterize marine pollution risk comes from the global
cumulative impacts assessment on habitats [@halpern2008b; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasives species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008b and @halpern2015.

@@

- Coastal development
- Direct human impact
- All fisheries data
- Hypoxia
- Inorganic pollution
- Invasive species
- Nutrient pollution
- Organic pollution
- Sea bottom temperature anomalies <!-- verify -->
- Shipping

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/marimekko.png}
\caption{Frequency distribution of cumulative exposure (*i.e.* sum of normalized driver intensity in each grid cell) and percent contribution of each driver to the frequency distribution of cumulative exposure in the Estuary and Gulf of St. Lawrence.}
\label{marimekko}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/validation.png}
\caption{Validation procedure for the *k-medoids* and *k-means* clustering algorithms based on the number of cluster that maximizes average silhouette width [upper panels; @kaufman1990] and minimizes the total within-cluster sum of squares (WSS; lower panels).}
\label{valid}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics{./Figures/clusterComp.png}
\caption{Comparison of the distribution of the 6 clusters identified using the *k-medoids* and *k-means* clustering algoriths and the percent grid cell correspondance between the two approaches.}
\label{comparison}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/interDissimilarity.png}
\caption{Evaluation of inter-cluster dissimilarity using a similarity percentage analysis (SIMPER) with Bray-Curtis dissimilarity [@clarke1993]. The figure diagonal presents the distribution of the 6 clusters identified using the *k-medoids* clustering algorithm. The lower triangle shows all combinations of inter-cluster dissimilarity with circular barplots showing the percent contribution to total dissimilarity of each driver and with the total inter-cluster dissimilarity in the center of the barplots. The upper triangle shows the average relative intensity of each driver for all driver combinations, with barplots to the left and the right representing the row and columns clusters, respectively.}
\label{inter}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/intraSimilarity.png}
\caption{Evaluation of intra-cluster similarity using the Bray-Curtis similarity index. The distribution of all 7 clusters is presented along with circular barplots showing the percent contribution to total similarity of each driver and with the total intra-cluster similarity in the center of the barplots.}
\label{intra}
\end{figure}

>>>>

<!--
cd phdproj/edrivers/edriversms
rmarkdown::render('./eDriversMS-SI-v2.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS-SI-v2.rmd', 'word_document')
-->

@@

Table S1. List of drivers currently available on *eDrivers* along with their respective acronym used in the figures in the supplementary material.

@@

We used $\Omega_{A}$ saturation state to characterize ocean acidification in the
bottom waters of the St. Lawrence. The data come from the Department of
Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018]
surveys and DFO's multispecies surveys [@bourdages2018] collected in
August-September of 2017 [@starr2019]. The carbonate chemistry was determined
through pH and TA measurements. Samples for pH and total alkalinity (TA) were
collected under a bubble free and no head space conditions into 500 $mL$
borosilicate glass flasks, 250 $\mu L$ of the saturated $HgCl_2$ solution was
added to sample, and processed following the "Guide to best practices for Ocean
CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurements using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The dilute
HCl titrant 0.1M in a solution of NaCl of 0.6M was calibrated using CRM provided
by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for in situ pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, total alkalinity, soluble reactive
and silicate concentrations as input parameters.

$\Omega_{A}$, pH, dissolved oxygen ($O_2$) and pressure were sampled at 117
stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The
majority of sampling stations were located in the Southern Gulf ($n = 84$),
compared to the Estuary and Northern Gulf ($n = 33$).

@@

We explored the correlation between $\Omega_{A}$ and $O_2$ in the data and found
no significant correlation at the scale of the whole St. Lawrence (*p-value*:
0.08; $R^2$: 0.03). Nevertheless, a correlation between $\Omega_A$ and $O_2$
exists in the Northern St. Lawrence (*p-value*: < 0.01; $R^2$: 0.49), but not
in the Southern St. Lawrence (*p-value*: 0.85; $R^2$: < 0.01). Furthermore, the
correlation observed in the north between $\Omega_A$ and $O_2$ is especially
strong when deep Channels (*Deep* model: *p-value*: < 0.01; $O_2$: 0.20; $R^2$:
0.97) are separated from shallower regions (*Shallow* model: *p-value*: < 0.01;
$O_2$: 0.07; $R^2$: 0.96), with $\Omega_A$ increasing more steeply with
increasing $O_2$ in deep Channels. This stronger correlation is likely caused
by the lack of gas exchange with the atmosphere or the surface mixed layer and
the importance of metabolic $CO_2$ in deep waters, some of which may have mean
ages of up to 20 years [@mucci2011]. In the Northern Gulf, the acidification is
following depth and distance from the Cabot Strait along the Laurentian,
Anticosti and Esquiman Channels due to the progressive oxygen depletion and
metabolic carbon dioxide accumulation [@mucci2011].

There are likely different dynamics at play in the Southern Gulf, such as
waters with lowered alkalinity and salinity caused by the St. Lawrence plume.
Such water masses are have lower buffering capacities to decreasing pH and
$\Omega_{A}$. Also, the accumulation of metabolic carbon dioxide produced by
rapid oxic degradation of organic matter in the sandy
and permeable sediments of the Magdalen Shelf or through the St. Lawrence plume
could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal
communication*).

@@

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentration in the water
column from the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a
covariable, as done in [@dutil2011]. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

@@

The data used to characterize hypoxia uses oxygen saturation in % and come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] surveys in late spring and fall, as well as from annual
DFO’s multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. We provide a brief summary of data and methods to
describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration is measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on the CTD. The probe is calibrated against seawater
samples that are analyzed by Winkler titration on every cast
[see @blais2019 for calibration procedure]. The data used is the last depth
sampled on the CTD profile, which typically corresponds to ~10m above the
bottom. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation taking into account the salinity and temperature at the selected
depth.

We interpolated oxygen saturation using cokriging with depth as a covariable,
as done in [@dutil2011]. Bathymetric data come from @dutil2011 and have a
100 $km^2$ resolution.   According to @diaz1995, severe hypoxia can be observed
when dissolved oxygen falls below 2 $ml$ $L^{-1}$, which corresponds to 62.5
$\mu mol$ $L^{-1}$ and ~20% saturation. This is considered as the level
necessary to maintain most animal life. Similarly, @chabot2008 studied the
effects of hypoxia on the energy budget of cod in the St. Lawrence and found
that behavioural effects began manifesting below 70% oxygen saturation and that
survival becomes jeopardized below ~20% saturation. We used these observations
to create an index of hypoxia using an inverted logistic curve that slowly
increases below 70% oxygen saturation, increases acutely as it approaches the
20% oxygen saturation threshold and reaches its peak intensity between 20 and
30% oxygen saturation:

@@

The surface layer is characterized using sea surface temperature (SST) monthly
composites from Advanced Very High Resolution Radiometer (AVHRR) satellite
images obtained from the National Oceanic and Atmospheric Administration (NOAA)
and European Organization for the Exploitation of Meteorological Satellites
(EUMETSAT). Images used are from DFO’s Maurice Lamontagne Institute at a 1 km
resolution from 1985-2013 and from DFO’s Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5 km resolution since 2014. Monthly
anomalies were constructed as the difference between monthly averages and the
1985-2010 climatological mean for each month.

@@

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level. We therefore had
to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence [@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].

@@

The data comes from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use globally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

@@

As in @halpern2008a and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statistics-canada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modelled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic pollution.

@@

The information provided on this expert map on the 5 toxins [@bates2019].
was georeferenced and transformed as vectorized objects.
We calculated a toxic algae risk ($T$) index for each cell ($x$) in the 1 $km^2$
study grid. For each toxin ($t$), a value of 1 was attributed to all grid cells
overlapping with areas identified at risk on the expert map and a value of 0.5
for grid cells overlapping with areas where ASP and DSP toxins were observed
without exceeding legal thresholds. The value for all 5 toxins was them summed
for all grid cells:

@@

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities in the
St. Lawrence. The data we used spans 6 years from 2010 to 2015 and details
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targeted species and a total of 53 caught species in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008a and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

@@

In order to characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008a; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

Two data sources were used to characterize shipping. The first set of data is
gathered as part of the World Meteorological Organization Voluntary Observing
Ships' (VOS) scheme. Ships participating in the program gather
meteorological data along with observation location as part of an open-ocean
climate dataset. The data spans 20 years and annually covers 10-20% of ships
worldwide. Data used spanned 2003 to 2011.

@@

The data used to characterize invasive species risk comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
supplementary materials provided in @halpern2008a and @halpern2015.

Cargo volume was used as a proxy of invasion risk under the assumption that
risk of invasion is proportional to tonnes of goods transferred through ports.
Cargo throughput in metric tonnes for the year 2011 was accessed through a
variety of sources [see supplementary material in @halpern2015 for more details]
and cross-matched with entries in the World Port Index database (WPI; available
from the National Geospatial-Intelligence Agency). A gap-filling procedure
using linear regression and sets of predictors related to port volume and
available in the WPI dataset was then applied to the WPI dataset to predict
missing cargo volume entries. Finally, volume data was distributed in marine
environments adjacent to ports using a diffusive plume model with an exponential
decay function that set the maximum spread distance to approximately 1000 km.
The plume model was then clipped to areas less than 60 m deep, as
invasive species are more likely to invade shallow areas.

@@

The data used to characterize marine pollution risk comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015].
Marine pollution was considered to be mainly driver by the shipping industry.
As such, the driver layer was constructed by combining the shipping (*i.e.*
shipping lanes) and invasive species (*i.e.* cargo volume) layers.
invasive. For more details, refer to supplementary materials provided in
@halpern2008a and @halpern2015.

@@

- Coastal development
- Direct human impact
- All fisheries data
- Hypoxia
- Inorganic pollution
- Invasive species
- Nutrient pollution
- Organic pollution
- Sea bottom temperature anomalies
- Shipping

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/marimekko.png}
\caption{Frequency distribution of cumulative exposure (\textit{i.e.} sum of normalized driver intensity in each grid cell) and percent contribution of each driver to the frequency distribution of cumulative exposure in the Estuary and Gulf of St. Lawrence.}
\label{marimekko}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/validation.png}
\caption{Validation procedure for the \textit{k-medoids} and \textit{k-means} clustering algorithms based on the number of cluster that maximizes average silhouette width (upper panels; Kaufman and Rousseeuw, 1990) and minimizes the total within-cluster sum of squares (WSS; lower panels).}
\label{valid}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/interDissimilarity.png}
\caption{Evaluation of inter-cluster dissimilarity using a similarity percentage analysis (SIMPER) with Manhattan distance (Clarke, 1993). The figure diagonal presents the distribution of the 6 clusters identified using the \textit{k-medoids} clustering algorithm. The lower triangle shows all combinations of inter-cluster dissimilarity with circular barplots showing the percent contribution to total dissimilarity of each driver and with the total inter-cluster dissimilarity in the center of the barplots. The upper triangle shows the average relative intensity of each driver for all driver combinations, with barplots to the left and the right representing the row and columns clusters, respectively.}
\label{inter}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics{./Figures/intraSimilarity.png}
\caption{Evaluation of intra-cluster similarity using the Manhattan distance transformed to a similarity index. The distribution of the 6 clusters is presented along with circular barplots showing the percent contribution to total similarity of each driver and with the total intra-cluster similarity in the center of the barplots.}
\label{intra}
\end{figure}
